Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators

### 2025-10-20 thesis/paper_draft.qmd
codes for plotting sim results (replacing the images)

path <- "D:/Github/Recon_Honours_Thesis"
knitr::opts_knit$set(root.dir = path)
# setwd(path)
```

@@

```{r load results}
path_sim <- "D:/Github/Recon_Honours_Analysis/sim/thesis_sim/"

sim1_1 <- readRDS("./thesis/lib/S4-2-1_T50_M500.rds")
sim1_2 <- readRDS("./thesis/lib/S4-2-1_T300_M500.rds")

sim2_1 <- readRDS("./thesis/lib/S36-6-1_T50_M500.rds")
sim2_2 <- readRDS("./thesis/lib/S36-6-1_T300_M500.rds")

sim_dense_50 <- readRDS("./thesis/lib/S100-10-3-1_T50_M200_dense.rds")
sim_dense_300 <- readRDS("./thesis/lib/S100-10-3-1_T300_M200_dense.rds")

sim_sparse_50 <- readRDS("./thesis/lib/S100-10-3-1_T50_M500_sparsebwgrp.rds")
sim_sparse_300 <- readRDS("./thesis/lib/S100-10-3-1_T300_M200_sparsebwgrp.rds")

sim3_1 <- sim_dense_50 # using same with next section simulation
sim3_2 <- sim_dense_300
```

@@

In our first set of experiments, we examine how MinT combined with the different estimators perform as the hierarchy expands. We generate synthetic data from the same VAR(1) framework described earlier, but vary the number of bottom‐level series, $n_b$, across two structures: a small structure with six groups of six bottom series ($n_b = 6$x$6 = 36$), and a much larger configuration with two groups of fifty ($n_b = 2$x$50 = 100$).

@@

![Percentage relative improvement in MSE of reconciled forecasts over the base forecasts in the 6 by 6 case (top row) and the 2 by 50 case (bottom row), T=50 (left column) and T=300 (right column), for 1- to 4-step-ahead forecasts. The positive (negative) entries indicate a decrease (increase) in MSE relative to base.](D:/Github/Recon_Honours_Thesis/figs/sim_MSE.png){#fig-sim-results-1}

@@

![Percentage relative improvement in Energy score in both hierarchies and time dimensions, for 1-step-ahead forecasts. The positive entries indicate a decrease in Energy score relative to base.](D:/Github/Recon_Honours_Thesis/figs/sim_energy.png){#fig-sim-results-2}

>>>>

path <- here::here()
knitr::opts_knit$set(root.dir = path)
# setwd(path)
```

@@

```{r load results}
path_lib <- paste0(path, "/thesis/lib/")

sim1_1 <- readRDS(paste0(path_lib, "S4-2-1_T50_M500.rds"))
sim1_2 <- readRDS(paste0(path_lib, "S4-2-1_T300_M500.rds"))

sim2_1 <- readRDS(paste0(path_lib, "S36-6-1_T50_M500.rds"))
sim2_2 <- readRDS(paste0(path_lib, "S36-6-1_T300_M500.rds"))

sim_dense_50 <- readRDS(paste0(path_lib, "S100-10-3-1_T50_M200_dense.rds"))
sim_dense_300 <- readRDS(paste0(path_lib, "S100-10-3-1_T300_M200_dense.rds"))

sim_sparse_50 <- readRDS(paste0(path_lib, "S100-10-3-1_T50_M500_sparsebwgrp.rds"))
sim_sparse_300 <- readRDS(paste0(path_lib, "S100-10-3-1_T300_M200_sparsebwgrp.rds"))

sim3_1 <- sim_dense_50 # using same with next section simulation
sim3_2 <- sim_dense_300

sim_MSE <- readRDS(paste0(path_lib, "sim_MSE.rds"))
sim_prob <- readRDS(paste0(path_lib, "sim_prob.rds"))
```

@@

In our main experiments, we examine how MinT combined with the different estimators perform as the hierarchy expands. We generate synthetic data from the same VAR(1) framework described earlier, but vary the number of bottom‐level series, $n_b$, across two structures: a small structure with six groups of six bottom series ($n_b = 6$x$6 = 36$), and a much larger configuration with two groups of fifty ($n_b = 2$x$50 = 100$).

@@

<!-- ![Percentage relative improvement in MSE of reconciled forecasts over the base forecasts in the 6 by 6 case (top row) and the 2 by 50 case (bottom row), T=50 (left column) and T=300 (right column), for 1- to 4-step-ahead forecasts. The positive (negative) entries indicate a decrease (increase) in MSE relative to base.](D:/Github/Recon_Honours_Thesis/figs/sim_MSE.png){#fig-sim-results-1} -->

```{r}
#| label: fig-sim-results-1
#| fig-cap: "Percentage relative improvement in MSE of reconciled forecasts over the base forecasts in the 6 by 6 case (top row) and the 2 by 50 case (bottom row), T=50 (left column) and T=300 (right column), for 1- to 4-step-ahead forecasts. The positive (negative) entries indicate a decrease (increase) in MSE relative to base. The negative improvements beyond 6% are capped."
#| fig.height: 10
#| fig.width: 10
#| out.width: "100%"

MSE_plots <- list()

cap <- -6
for (sim_name in names(sim_MSE)) {

  plot_df <- sim_MSE[[sim_name]] |> 
    # cap pct_change upper 10%
    mutate(
      capped = pct_change <= cap,
      capped_pct_change = pmax(pct_change, cap)
    ) |> 
    mutate(shape_lab = ifelse(shape_lab %in% c("PC1","PC2"), shape_lab, ""))
  p <- plot_df |>
    ggplot(aes(
      x = h, y = capped_pct_change,
      color = family, shape = shape_lab, linetype = variant_type
    )) +
      geom_line(linewidth = 0.8) +
      geom_point(alpha = 0.85, size = 2.6, na.rm = TRUE) +
      recon_scales +
      # # label actual values for capped points
      # geom_text(
      #   data = subset(plot_df, pct_change <= cap),
      #   aes(label = round(pct_change, 1)),
      #   vjust = -0.5, color = "black", size = 3
      # ) +
      theme_minimal() +
      labs(x = "Forecast Horizon", y = "% improvements")
  MSE_plots[[sim_name]] <- p
}

MSE_plots[[1]] + labs(title = "6 Groups of 6", subtitle = "T=50") +
  MSE_plots[[2]] + labs(subtitle = "T=300") +
  MSE_plots[[3]] + labs(title = "2 Groups of 50", subtitle = "T=50") +
  MSE_plots[[4]] + labs(subtitle = "T=300") +
  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") & 
  theme(legend.position = "bottom", legend.direction = "vertical", legend.title.position = "left")
```

@@

<!-- ![Percentage relative improvement in Energy score in both hierarchies and time dimensions, for 1-step-ahead forecasts. The positive entries indicate a decrease in Energy score relative to base.](D:/Github/Recon_Honours_Thesis/figs/sim_energy.png){#fig-sim-results-2} -->

```{r}
#| label: fig-sim-results-2
#| fig-cap: "Percentage relative improvement in Energy score in both hierarchies and time dimensions, for 1-step-ahead forecasts. The positive entries indicate a decrease in Energy score relative to base."
#| fig.height: 10
#| fig.width: 10
#| out.width: "100%"

energy_plots <- list()

for (sim_name in names(sim_prob)) {

  p <- sim_prob[[sim_name]] |>
    filter(rule == "Energy") |>
    filter(model_label != "Base") |>
    ggplot(aes(x = model_label, y = pct_change_avg, fill = family, alpha=pc_k)) +
      geom_col(color = "black") +
      scale_alpha_manual(values = c("0" = 1, "1" = 0.7, "2" = 0.4), guide = "none") +
      recon_fill_scale +
      labs(x = "Model", y = "% improvements") +
      theme_minimal() +
      theme(
        axis.text.x = element_text(angle = 20, hjust = 0.5)
      )

  energy_plots[[sim_name]] <- p
}

energy_plots[[1]] + labs(title = "6 Groups of 6", subtitle = "T=50") +
  energy_plots[[2]] + labs(subtitle = "T=300") +
  energy_plots[[3]] + labs(title = "2 Groups of 50", subtitle = "T=50") +
  energy_plots[[4]] + labs(subtitle = "T=300") +
  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") & 
  theme(legend.position = "bottom", legend.title = element_blank())

```
